---
title: "Prediction Assignment"
author: "Osama Mohsen"
date: "Augest 2, 2017"
output: html_document
---
### Load libraries
```{r}
library(caret)
```

### load data
* Load train and test datasets.           
```{r}
if(!file.exists("pml-training.csv")){
        trainDataURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
        download.file(trainDataURL, destfile = "pml-training.csv")
        }
if(!file.exists("pml-testing.csv")){
        testDataURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
        download.file(testDataURL, destfile = "pml-testing.csv")
        }

trainData <- read.csv("pml-training.csv")
testData <- read.csv("pml-testing.csv")
```
### analysis the datasets
```{r}
names(trainData)
summary(trainData)
str(trainData)
dim(trainData)
colSums(is.na(trainData))/dim(trainData)[1]
```

### clean the dataset
based on the analysis we observe alot of entries with na values so remove ones with na values more than 97%.
```{r}
naCols  <- colSums(is.na(trainData))
trainData <- trainData[ ,naCols/dim(trainData)[1] < 0.97]
```
### data preprocessing

we remove features with zerovariance as it will not affect our decision boundry and training process.

```{r}
zerovar <- nearZeroVar(trainData, saveMetrics = FALSE)
zerovar
trainData <- trainData[ , -zerovar]
nearZeroVar(trainData, saveMetrics = FALSE)
dim(trainData)
```

### cross validation
we split the training dataset to 75% for training and 25% to validation and preformance measure.
```{r}
inTrain <- createDataPartition(trainData$classe, p = 0.70, list = FALSE)
trainPart <- trainData[inTrain, ]
validatePart <- trainData[-inTrain, ]
```

### cross validation
we split the training dataset to 75% for training and 25% to validation and preformance measure. we als 3-fold cross-validation.
```{r data, eval = F}
inTrain <- createDataPartition(trainData$classe, p = 0.70, list = FALSE)
trainPart <- trainData[inTrain, ]
validatePart <- trainData[-inTrain, ]
```

### training and model selection
#### Model Selection:
The problem goal is is classification. We would use random forest as a model and we would measure it preformance although
it might take long in training phase it would achieve high accurecy. we use 3-fold cross-validatio for our training.
```{r}
set.seed(111)
trainCont <- trainControl(method="cv",number=3)
forestModel = train(classe ~., data=trainPart, method="rf",trControl=trainCont)
forestModel
```
### validation and preformance measure
```{r}
validationMeasure <- predict(forestModel, validatePart)
table(validationMeasure, validatePart$classe)
confusionMatrix(validatePart$classe, validationMeasure)

modelURL <- "forestModel.RData"
save(forestModel, file=modelURL )
```
We acieved accurecy of 1.0 on both training and validation datasets.
### testing and predection
now we use your prediction model to predict 20 different test cases.
```{r}
testMeasure <- predict(forestModel, newdata = testData)
testMeasure
```